---
title: 'Class 15: RNASeq analysis'
author: 'Barry (PID: 911)'
date: "11/17/2021"
output: html_document
---

# Background

Today we examine a published RNA-seq experiment where airway smooth muscle cells were treated with dexamethasone, a synthetic glucocorticoid steroid with anti-inflammatory effects (Himes et al. 2014).

## Load the contData and colData

We need 2 things 
- 1: count data
- 2: colData (the metadata that tells us about the design of the experiment).


```{r}
counts <- read.csv("airway_scaledcounts.csv", row.names=1)
metadata <- read.csv("airway_metadata.csv")
```

```{r}
head(counts)
```

```{r}
head(metadata)
```

Side-note:
Let's check the corespondence of the metadata and count data setup.

```{r}
metadata$id
```

```{r}
colnames(counts)
```

We can use the `==` thing to see if they are the same

```{r}
all(metadata$id == colnames(counts))

```

```{r}
all( c(T,T,T,T,T,T,F) )
```

## Compare control to treated

First wee need to access all the control columns in our counts data.

```{r}
control.inds <- metadata$dex == "control"
control.ids <- metadata[ control.inds, ]$id
```

Use these ids to access just the control columns of our `counts` data

```{r}
control.mean <- rowMeans(counts[ , control.ids] )
head(control.mean)
```

Do the same for the drug treated...
```{r}
treated.id <- metadata[ metadata$dex == "treated", ]$id
treated.mean <- rowMeans(counts[,treated.id])
```


We will combine our meancount data for bookkeeping purposes.

```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

There are `r nrow(counts)` rows/genes in this dataset

```{r}
nrow(counts)
```

## Compare the control and treated

A quick plot of our progress so far

```{r}
plot(meancounts)
```

This would benefit from a log transform! Let's plot on a log log scale

```{r}
plot(meancounts, log="xy")
```

We often use log transformations as they make life much nicer in this world...

```{r}
log2(20/20)
```
```{r}
log2(40/20)
```
```{r}
log2(10/20)
```
```{r}
log2(80/20)
```

Cool. I like log2!

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/
                          meancounts[,"control.mean"])
head(meancounts)
```

We need to drop the zero count genes/rows!

```{r}
head(meancounts[,1:2])
```


```{r}
head(meancounts[,1:2] == 0)
```

The `which()` function tells us the indices of TRUE entries in a logical vector.

```{r}
which( c(T,F,T) )
```
However, it is not that useful in default mode on our type of multi column input....

```{r}
inds <- which(meancounts[,1:2] == 0, arr.ind=TRUE)
head(inds)
```
I only care about the rows here (if there is a zero in any column I will exclude this row eventually).

```{r}
to.rm <- unique(sort(inds[,"row"]))
```

```{r}
mycounts <- meancounts[-to.rm,]
head(mycounts )
```

We now have `r nrow(mycounts)` genes remaining.

```{r}
nrow(mycounts)
```
How many of these genes are up regulated at the log2 fold-change threshold of +2 or greater?

```{r}
sum(mycounts$log2fc > +2)
```

What percentage is this?

```{r}
round((sum(mycounts$log2fc > +2) / nrow(mycounts))*100, 2)
```
How about down?

```{r}
sum(mycounts < -2)
```

# DESeq2 analysis

```{r}
library(DESeq2)
```
We first need to setup the DESeq input object.

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex)
```
Run the DESeq analysis pipeline.

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
head(res)
```

# A Volcano plot

This is a very common data viz of this type of data that does not really look like a volcano.

```{r}
plot(res$log2FoldChange, -log(res$padj))
abline(v=c(-2,2), col="gray")
abline(h=-log(0.05), col="gray" )
```

## Adding annotation data

We want to add meaningful gene names to our dataset so we can make some sense of what is going on here...

For this we will use two bioconductor packages, one does the work and is called **AnnotationDbi** the other contains the data we are going to map between and is called **org.Hs.eg.db**.

First I need to install with `BiocManager::install("org.Hs.eg.db")` and
`BiocManager::install("AnnotationDbi")`.

```{r}
library(AnnotationDbi)
library(org.Hs.eg.db)
```
```{r}
columns(org.Hs.eg.db)
```

Here we map to "SYMBOL" the comon gene name that the world understands and wants.

```{r}
res$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     keytype="ENSEMBL",   
                     column="SYMBOL",   
                     multiVals="first")

```


```{r}
head(res)
```

# Lets finally save our results to date

```{r}
write.csv(res, file = "allmyresults.csv")
```

