---
title: 'Class 10: Candy mini project'
author: 'Barry (PID: 911)'
date: "10/29/2021"
output: github_document
---

# Read the data

This comes from the 538 GitHub repo. They have lot's of interesting datasets.

```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy <- read.csv(url, row.names=1)
head(candy)
```
```{r}
gsub("Õ", "'", rownames(candy))
```
> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```
```{r}
sum(candy$chocolate)
```


```{r}
library(skimr)
skim(candy)
```

```{r}
hist(candy$winpercent)
```

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
choclate <- candy[ as.logical(candy$chocolate), ]$winpercent
mean(choclate)
```
```{r}
fruity <- candy[as.logical(candy$fruity),]$winpercent
mean(fruity)
```


> Q12. Is this difference statistically significant

Yes!

```{r}
t.test(choclate, fruity)
```

# 3. Candy rankings
Let's make a barplot of the winpercent values for the variouys candy types

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, rownames(candy)) +
  geom_col()

```

We need to improve this to reorder the candy by the winpercent valeus

```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```

## Time to add some color

```{r}
# Color vector (all black to start)
my_cols=rep("black", nrow(candy))

# Now overwrite the chocolate entries with "chocolate"
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
my_cols
```


```{r}
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col(fill=my_cols)
```


```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) +
  geom_text()
```

These labels suck. Let's use the **ggrepl** package for better label placment.

```{r}
library(ggrepel)

# change my fruity color to red
my_cols[as.logical(candy$fruity)] <- "red"

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) +
  geom_text_repel(col=my_cols)

```


```{r}
rownames(candy) <- gsub("Õ", "'", rownames(candy))
```

```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) +
  geom_text_repel(col=my_cols, max.overlaps = 7)

```

# Correlation Analysis

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)

```

# Principla Componet analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
# PCA score plot

```{r}
plot(pca$x[,1:2], col=my_cols)
```

ggplot version

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])

ggplot(my_data) +
  aes(PC1, PC2) +
  geom_point(col=my_cols)
```

