---
title: "Machine Learning 1"
author: 'Barry (PID: 911)'
date: "10/22/2021"
output: github_document
---

# Clustering methods

Kmeans clustering in R is done with the `kmeans()` function.
Here we makeup some data to test and learn with.

```{r}
tmp <- c(rnorm(30, 3), rnorm(30, -3) )
data <- cbind(x=tmp, y=rev(tmp))
plot(data)
```

Run `kmeans()` set k (centers) to 2 (i.e. the number of clusters we want) nstart 20 (to tun multiple times). The thing with Kmeans is you have to tell it how many clusters you want.

```{r}
km <- kmeans(data, centers = 2, nstart=20)
km
```

> Q. How many points are in each cluster?

```{r}
km$size
```
> Q. What ‘component’ of your result object details cluster assignment/membership? 

```{r}
km$cluster
```

> Q. What ‘component’ of your result object details cluster center?

```{r}
km$centers
```

> Q. Plot x colored by the kmeans cluster assignment and add cluster centers as blue points

```{r}
plot(data, col=km$cluster)
points(km$centers, col="blue", pch=15, cex=2)
```

# Hierarchical Clustering

We will use the `hclust()` function on the same data as before and see how this method works.

```{r}
hc <- hclust( dist(data) )
hc
```

hclust has a plot method

```{r}
plot(hc)
abline(h=7, col="red")
```

To find our membership vector we need to "cut" the tree and for this we use the `cutree()` function and tell it the height to cut at.

```{r}
cutree(hc, h=7)
```

We can also use `cutree()` and sate the number of k clusters we want...

```{r}
grps <- cutree(hc, k=2)
```

```{r}
plot(data, col=grps)
```


# Principal Component Analysis (PCA)

PCA is a super useful analysis method when you have lots of dimensions in your data...

## PCA of UK food data

Import the data from a CSV file

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
```

How many rows and cols?
```{r}
dim(x)
```
```{r}
x[,-1]
```
```{r}
rownames(x) <- x[,1]
x <- x[,-1]
x
```

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names=1)
x
```

```{r}
barplot( as.matrix(x), col=rainbow(17) )
```


```{r}
barplot( as.matrix(x), col=rainbow(17), beside=TRUE )
```

```{r}
mycols <- rainbow( nrow(x) )
pairs(x, col=mycols, pch=16)
```

## PCA to the rescue!

Here we will use the base R function for PCA, which is called `prcomp()`. This function wants the transpose of of data.

```{r}

pca <- prcomp( t(x) )
summary(pca)
```

```{r}
plot(pca)
```

We want score plot (a.k.a. PCA plot). Basically of PC1 vs PC2

```{r}
attributes(pca)
```

We are after the pca$x component for this plot...

```{r}
plot(pca$x[,1:2])
text(pca$x[,1:2], labels = colnames(x))
```

```{r}
library(ggplot2)
df <- as.data.frame(pca$x)

ggplot(df) + 
  aes(x=PC1, y=PC2, label=rownames(df)) +
  geom_text()
```



We can also examine the PCA "loadings", which tell us how much the original variables contribute to each new PC...

```{r}
par(mar=c(10, 3, 0.35, 0))

barplot(pca$rotation[,1], las=2)
```


## One more PCA for today

```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```

```{r}
nrow(rna.data)
```

```{r}
ncol(rna.data)
```
```{r}
colnames(rna.data)
```
```{r}
pca.rna <- prcomp( t(rna.data), scale=TRUE )
summary(pca.rna)
```

```{r}
plot(pca.rna)
```

```{r}
plot(pca.rna$x[,1:2])
text(pca.rna$x[,1:2], labels = colnames(rna.data))
```

Let's Knit to PDF and also github_document then push to github.

