---
title: 'Class17: COVID-19 vaccination rate mini project'
author: 'Barry (PID: 911)'
date: "11/24/2021"
output:
  pdf_document:
    latex_engine: xelatex
---

# Background

In this begore Thanksgiving class when many of our class mates are traveling let's have a look at COVID-19 vaccination rates around the State.

We get vaccination rate data from CA.GOV here:

https://data.ca.gov/dataset/covid-19-vaccine-progress-dashboard-data-by-zip-code

## Import data

```{r}
vax <- read.csv("covid19vaccinesbyzipcode.csv")
head(vax)
```


> Q. How many entries do we have?

```{r}
nrow(vax)
```

We can use the **skimr** package and the `skim()` function to get a quick overview of structure of this dataset.

```{r}
skimr::skim(vax)
```

Notice that one of these columns is a date column. Working with time and dates get's annoying quickly. We can use the **lubridate** package to make this easy...

```{r}
library(lubridate)

today()
```


> Q. How many days since the first entry in the dataset?

```{r}
vax$as_of_date[1]
```

This will not work because our data column was read as character..
```{r}
# today() - vax$as_of_date[1]
```

```{r}
d <- ymd(vax$as_of_date)
```

```{r}
today() - d[1]
```

I will make the `as_of_date` coulumn Date format...

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```


> Q. When was the dataset last updated? What it is the last date in this dataset? How many days since the last update?

```{r}
today() - vax$as_of_date[ nrow(vax) ]
```
> Q. How many days does the dateset span?

```{r}
vax$as_of_date[ nrow(vax) ] - vax$as_of_date[1]
```

> Q. How many different ZIP code areas are in this dataset?

```{r}
length(unique(vax$zip_code_tabulation_area))
```

To work with ZIP codes we can use the **zipcodeR**

```{r}
library(zipcodeR)

reverse_zipcode(c('92037', "92109") )
```

## Focus in on San Diego County

We want to subset the full CA `vax` data down to just San Diego County.

We could do this with base R

```{r}
inds <- vax$county == "San Diego"
nrow(vax[inds,])
```

Subsetting can get tedious and complicated quickly when you have multiple things we want to subset by.

```{r}
library(dplyr)
```

We will use the `filter()` function to do our subsetting from now on.

We want to focus in on San Diego County

```{r}
sd <- filter(vax, county=="San Diego")
nrow(sd)
```

More complicated subsetting...

```{r}

sd.20 <- filter(vax, county=="San Diego",
       age5_plus_population > 20000)

nrow(sd.20)
```

> Q. What is the average vaccination rate of San Diego count as of yesterday?

```{r}
sd.now <- filter(vax, county=="San Diego",
                 as_of_date=="2021-11-23")

head(sd.now)
```


```{r}
summary(sd.now$percent_of_population_fully_vaccinated)
```



> Q. Make a histogram of these values.

Base R histogram
```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```

This plot above is going to be susceptible to being skewed by ZIP code areas with small populations. Thes will have big effects for just a small number of unvax-ed folks...

> Q. What is the population of the 92037 ZIP code area?

```{r}
lj <- filter(sd.now, zip_code_tabulation_area=="92037")
lj$age5_plus_population
```

> Q. What is the average vaccination value for this UCSD/La Jolla ZIP code area?

```{r}
lj$percent_of_population_fully_vaccinated
```

> Q. What about this ZIP code 92122

```{r}
lj2 <- filter(sd.now, zip_code_tabulation_area=="92122")
lj2$age5_plus_population
```
```{r}
lj2$percent_of_population_fully_vaccinated
```

```{r}
filter(sd.now, zip_code_tabulation_area=="92124")
```

---

Time series of vaccination rate for a given ZIP code area.
Start with 92037.

```{r}
lj <- filter(vax, zip_code_tabulation_area=="92037")
```

Use ggplot for this:
```{r}
library(ggplot2)

ggplot(lj) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  labs(x="Date", y="Precent Vaccinated")

```


Let's make this plot for all San Diego County ZIP code areas that have a population as least as large as 92037.

```{r}

sd.36 <- filter(vax, county=="San Diego",
             age5_plus_population > 36144)

head(sd.36)
```

> How many ZIP code areas in San Diego county have a population larger than 92037

```{r}
length(unique(sd.36$zip_code_tabulation_area))
```

Lets make the plot

```{r}
ggplot(sd.36) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area) +
  geom_line() +
  labs(x="Date", y="Precent Vaccinated")

```
> Q. Make a plot like this for the all ZIP code areas in the State with a population at least as large as La Jolla.

```{r}
ca <- filter(vax, age5_plus_population > 36144)
```

How many
```{r}
length(unique(ca$zip_code_tabulation_area))
```
Make our big monster plot

```{r}
ggplot(ca) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2) +
  labs(x="Date", y="Precent Vaccinated") +
  geom_hline(yintercept = 0.67, color="red")

```

> Q. What is the mean across the state for these 36k + population areas?

```{r}
ca.now <- filter(ca, as_of_date=="2021-11-23")
summary( ca.now$percent_of_population_fully_vaccinated )
```






